PHQ-4-R - Data Cleaning (Study 2)

### 2024-05-21 study1/analysis.qmd
update

```{r}
df_anx <- select(df, contains("Anxiety"))
df_anx2 <- select(df2, contains("Anxiety"))
```

@@

@@

## Model

@@

## Category Characteristic Curves (CRC) and Item Information Curves

@@

## Normalized scoring

>>>>

```{r}
df_anx <- select(df, contains("Anxiety"))
df_anx2 <- select(df2, contains("Anxiety"))
df_anxfull <- rbind(df_anx, df_anx2)
```

@@

# Plots
plot(m_anxiety2, type = "trace", theta_lim = c(-3, 3))
plot(m_anxiety2, type = "infotrace", theta_lim = c(-3, 3))
plot(m_anxiety2, type = "score", theta_lim = c(-3, 3))
plot(m_anxiety2, type = "infoSE", theta_lim = c(-3, 3))
plot(m_anxiety2, type = "rxx", theta_lim = c(-3, 3))
```

#### Full

```{r}
m_anxietyfull <- mirt(df_anxfull, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

plot(m_anxietyfull, type = "trace", theta_lim = c(-3, 3))
# plot(m_anxietyfull, type = "infotrace", theta_lim = c(-3, 3))
```

@@

@@

@@

### 2024-05-21 study2/analysis/1_cleaning.qmd
update

@@

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
  geom_point(aes(color = Experimenter, shape = Category)) +
  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  guides(shape = "none") +
  ggside::geom_xsidedensity(aes(fill=Experimenter), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_bw() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) 
```

@@

# df |> 
#   ggplot(aes(x=BDI2_Duration, y=BDI2_SD)) +
#   geom_point()

table <- df |> 
  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
  mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
  mutate(Participant = fct_reorder(Participant, outliers$Duration)) |> 
  arrange(Participant)

data.frame(Participant = c("Average"), t(sapply(table[2:ncol(table)], mean, na.rm = TRUE))) |>
  rbind(table) |> 
  gt::gt() |> 
  gt::fmt_number() |> 
  # gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers_half, "orange", "white")) |>
  gt::data_color(columns = ends_with("Duration"), 
                 direction="column", palette = "RdBu") |> 
  gt::data_color(columns = ends_with("SD"), 
                 direction="column", palette = "RdBu", reverse=TRUE) |> 
  gt::data_color(direction="row", rows=1, palette = "grey") |> 
  gt::opt_interactive(use_compact_mode = TRUE)
```

### Attentional Checks

```{r}
df |> 
  select(starts_with("IAS_"), -IAS_Duration, -IAS_Order) |>
  pivot_longer(starts_with("IAS_"), names_to = "item", values_to = "score") |> 
  ggplot(aes(x = score)) +
  geom_density() +
  facet_wrap(~item)
```

@@

```{r}
add_bf <- function(r) {
  x <- as.data.frame(r)
  f <- as.formula(paste(str_remove(x$Parameter, fixed("df$")), "~ PHQ4_Condition"))
  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
  x
}

rbind(
  add_bf(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
) |> 
  select(Parameter, t, df_error, p, BF) |> 
  mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
  insight::display()
```

@@

`r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment. `r sum(df$DisorderHistory)` participants (`r insight::format_percent(sum(df$DisorderHistory) / nrow(df))`) reported having a history of mood disorder.

>>>>

df$Sample <- case_when(
  df$Experimenter %in% c("Experimenter1", "Experimenter2") ~ "Student",
  df$Experimenter %in% c("dom", "Unknown", "readme") ~ "Other",
  .default = df$Experimenter
)
```

@@

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
  geom_point(aes(color = Sample, shape = Category)) +
  geom_text(aes(color = Sample, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  guides(shape = "none") +
  ggside::geom_xsidedensity(aes(fill=Sample), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_bw() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) 
```

### Attention Checks

Only a subset of participants saw the version with attention checks.

```{r}
reject <- df |> 
  filter(!is.na(AttentionCheck_2)) |>
  select(Participant, Experimenter, starts_with("AttentionCheck")) |> 
  arrange(AttentionCheck_2, AttentionCheck_3)

failed <- reject[reject$AttentionCheck_2 == 0 | reject$AttentionCheck_3 < 0.7, "Participant"]
failed <- failed[!is.na(failed)]

reject |> 
  gt::gt() |> 
  gt::data_color(columns = c("AttentionCheck_2", "AttentionCheck_3"), palette=c("red", "green")) |> 
  gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% failed, "red", "white")) |> 
  gt::opt_interactive(use_compact_mode = TRUE)

df <- filter(df, !Participant %in% failed)
df <- select(df, -starts_with("AttentionCheck"))
```

We dropped `r length(failed)` participants for failing the attention checks.

@@

mv_outliers <- df |> 
  select(Participant, 
         matches("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total,
         matches("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General) |> 
  check_outliers(method="mahalanobis_robust", ID="Participant")

df |>
  mutate(Outliers = mv_outliers) |> 
  ggplot(aes(x=STAI5_SD, y=BDI2_SD)) +
  geom_point(aes(color=Sample, shape=mv_outliers), alpha=2/3, size=3) +
  scale_shape_manual(values=c(16, 4)) +
  ggside::geom_xsidedensity(aes(color=Sample), alpha=0.5) +
  ggside::geom_ysidedensity(aes(color=Sample), alpha=0.5) +
  theme_minimal() +
  ggside::theme_ggside_void()

table <- df |> 
  select(Participant, starts_with("AttentionCheck"), BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
  mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
  mutate(Participant = fct_reorder(Participant, outliers$Duration)) |> 
  arrange(Participant)

data.frame(Participant = c("Average"), t(sapply(table[2:ncol(table)], mean, na.rm = TRUE))) |>
  rbind(table) |> 
  gt::gt() |> 
  gt::fmt_number() |> 
  # gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers_half, "orange", "white")) |>
  gt::data_color(columns = ends_with("Duration"), 
                 direction="column", palette = "RdBu") |> 
  gt::data_color(columns = ends_with("SD"), 
                 direction="column", palette = "RdBu", reverse=TRUE) |> 
  gt::data_color(direction="row", rows=1, palette = "grey") |> 
  gt::opt_interactive(use_compact_mode = TRUE)

df <- df[!mv_outliers,]
```

We dropped `r length(mv_outliers)` participants flagged as multivariate outliers using the robust Mahalanobis distance.

@@

```{r}
add_bf <- function(df, var="Age") {
  r <- report::report(t.test(df[[var]] ~ df[["PHQ4_Condition"]]), verbose=FALSE)
  x <- as.data.frame(r)
  f <- as.formula(paste(var, "~ PHQ4_Condition"))
  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
  x$Parameter <- var
  x
}

rbind(
  add_bf(filter(df, !is.na(Age)), "Age"),
  add_bf(df, var="PHQ4_Total"),
  add_bf(df, "STAI5_General"),
  add_bf(df, "BDI2_Total")
) |> 
  select(Parameter, t, df_error, p, BF) |> 
  mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
  insight::display()
```

@@

`r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment. `r sum(df$DisorderHistory, na.rm=TRUE)` participants (`r insight::format_percent(sum(df$DisorderHistory, na.rm=TRUE) / nrow(df))`) reported having a history of mood disorder.
